First Season MWA EoR Power Spectrum Results at Redshift 7

### 2016-08-01 ms.tex
Updated Carroll et al 2016 reference
Fixed citealt->citet

Several probes are being used to investigate the EoR. 
Studies of the polarization of the CMB have placed integrated constraints on 
the timing of reionization \citep[e.g.][]{Planck:2015b, Planck:2015, WMAP9}. Meanwhile, 
observations of highly redshifted quasars have placed upper bounds on redshifts by which 
reionization is complete. For example, \citealt{Fan:2006a} showed that reionization must be 
complete by $z \approx 6$, a result which was confirmed independent of model by 
\citealt{McGreer:2015}.
Using a $z_{\mathrm{end}}>6$ prior, the most recent analysis of \emph{Planck} data 
found a reionization redshift $z_{\mathrm{re}} = 8.8 \pm 0.9$ for a redshift-symmetric model
($z_{\mathrm{re}} = 8.5 \pm 0.9$ for redshift-asymmetric), and a duration of $\Delta z < 2.8$
\citep{Planck:2016}.
Deep optical and infrared galaxy surveys are also beginning to 
reach the redshifts necessary to further constrain reionization \citep[e.g.][]{Bouwens:2014}, 
and the James Webb Space Telescope will improve on their sensitivity 
\citep{Gardner:2006}.

@@

Using the first generation of instruments, several upper limits have been placed on the 21 
cm EoR signal \citep{Ali:2015, Dillon:2014, Parsons:2014, Jacobs:2015, Paciga:2013}, and 
\citealt{Pober:2015} was able to place constraints on physical reionization models. 
However, much work is to be done to understand the data produced by these arrays. 
Foreground subtraction and isolation has emerged as a priority in the field. Central to most 
analysis strategies is the concept of an ``EoR window" -- the region of Fourier space where 
spectrally smooth foregrounds have been isolated from the isotropic (spherically symmetric 
in Fourier space) cosmological signal \citep{Morales:2006, Bowman:2009}. More recent 
studies have shown the existence of a foreground ``wedge", the result of instrumental mode 
mixing, throwing power from spectrally smooth foregrounds to higher Fourier modes 
\citep{Thyagarajan:2015b, Thyagarajan:2015, Trott:2012, Liu:2014a, Liu:2014b, 
Hazelton:2013, Pober:2013, Thyagarajan:2013, Vedantham:2012, Morales:2012, Datta:2010}, 
while others are investigating the spectral structure of point sources themselves \cite[e.g.][]{Offringa:2016}.
The EoR window is still expected to be preserved above the wedge, and 
several analysis pipelines are under active development \citep[e.g. B.~J.~Hazelton et al. 2016, in 
prep; D.~A.~Mitchell et al. 2016, in prep; ][]{Jacobs:2016, Trott:2016, 
Dillon:2015b, Dillon:2013, Trott:2014}.

@@

Our best result is a limit on the cosmological power spectrum of $\Delta^2 \leq 2.7 \times 10^4$ 
mK$^2$ at $k=0.27$ h~Mpc$^{-1}$ and $z=7.1$. This is lower than the GMRT 40 hour
limit ($\Delta^2 \le 6.15 \times 10^4$ mK$^2$ at $z=8.6$; \citealt{Paciga:2013}), but 
signficantly higher than the latest PAPER result which 
integrated over 700 hours ($\Delta^2 \le 502$ mK$^2$ at $z=8.4$; \citealt{Ali:2015}).
Several EoR experiments are reaching sensitivity levels where very low systematics
become dominant. The PAPER team saw evidence that antenna cross-talk and foreground
leakage was limiting their result at several scales. The LOFAR team has seen excess noise
and diffuse foreground suppression due to calibration \citep{Patil:2016}. It is essential
for EoR experiments to understand and overcome these systematics in order to reach
a cosmological detection.

@@

\section{The Murchison Widefield Array and Observations}\label{sec:MWA}
The Murchison Widefield Array is one of several first generation radio interferometers with 
a primary science goal of detecting the 21cm EoR power spectrum. The remote Australian 
outback offers relative isolation from human-made radio frequency interference (RFI) such as 
FM radio, or TV stations. A recent study of the RFI environment at the Murchison Radio 
Observatory can be found in \citealt{Offringa:2015}.

While the 21 cm EoR signal is priority, the MWA is a general observatory serving several 
science programs beyond cosmology including Galactic and extragalactic surveys, time 
domain astrophysics, solar monitoring, and ionospheric science. The array was thus 
designed with these programs in consideration and the layout was optimized using a 
pseudo-random antenna placement algorithm \citep{Beardsley:2012} to obtain a well 
behaved point spread function while retaining a dense core for EoR sensitivity 
\citep{Beardsley:2013, Bowman:2006}.\footnote{A drone flyover view of the array layout is available on YouTube at \url{https://youtu.be/yDWdTUzTUMo}.}
A full description of the science capabilities of the array is found in
\citealt{Bowman:2013}.
The high imaging capability of the MWA enables us to calibrate and subtract foregrounds
based on sky models, leveraging the full capability of the observatory for the
EoR experiment. 
This strategy is in contrast to more targeted EoR experiments like PAPER, 
which utilize a highly redundant layout to enhance sensitivity. Redundant arrays can 
exploit symmetries of the instrument for quick calibration and analysis, but have poor 
point spread functions, making foregrounds more difficult to characterize.

The technical design of the MWA is reviewed in \citealt{Tingay:2013}, and we highlight a 
few key characteristics that will become important in our analysis of the data. Each 
antenna of the MWA comprises 16 dual-polarization dipoles placed on a regular grid, lying 
on a ground screen. The radio frequency signals from these dipoles feed into an analog 
beamformer which uses physical delays to ``point" the antenna. The MWA contains 128 of 
these antennas, with a tightly packed core of radius 50 m, and extending out to a radius of
1.5 km for higher resolution calibration and imaging.

@@

\section{Analysis Pipeline}\label{sec:analysis}
In order to ensure consistency in analysis, the international MWA EoR collaboration has 
defined two independent reference analysis pipelines. The details of this strategy are 
outlined in \citealt{Jacobs:2016}. This study is based on the FHD-to-\eppsilon 
pipe, which is in turn based on the Fast Holographic Deconvolution 
(FHD\footnote{\url{https://github.com/EoRImaging/FHD}}, \citealt{Sullivan:2012}) and Error 
Propagated Power Spectrum with Interleaved Observed Noise (\eppsilon
\footnote{\url{https://github.com/EoRImaging/eppsilon}}, B.~J.~Hazelton et al. 2016, in 
prep) packages. The general flow of the pipeline is shown in Figure~\ref{fig:pipe}, and we 
give an overview below. As a point of comparison, and to demonstrate robustness of the results, we compute the output power spectra with an independent calibration and power spectrum estimation pipeline, and demonstrate consistency.

@@

Calibration is accomplished within the FHD package in two steps. First we match the raw 
data to model visibilities formed using a realistic simulation of the array response to our 
foreground model. This step results in independent complex gains for each antenna, 
80~kHz frequency channel, and polarization. In the second step we impose restrictions on these 
gains motivated by our understanding of the spectral response of the instrument. By 
reducing the number of free parameters in this step we increase our signal to noise on 
calibration solutions and avoid absorbing unmodeled confusion source flux density into our gain 
solutions which has been shown to contaminate the EoR window \citep{Barry:2016}. 
Recently, \citealt{Patil:2016} demonstrated excess noise and a suppression of diffuse
foregrounds in LOFAR images after calibration and subtraction of bright discrete
foreground sources. They suggested a multi-frequency calibration solution as a potential
solution to these systematics, similar to the scheme implemented here. Indeed we have
not seen evidence of these systematics through simulations and direct propagation of
our noise \citep[Hazelton et al. 2016, in prep.]{Barry:2016}.

@@

FHD utilizes the StEFCal algorithm described in \citealt{Salvini:2014} to find the minimum 
$\chi^2$ estimate with respect to the complex gains. The result of this operation is 
estimated gains for every antenna, frequency channel, and polarization. However, with 
certain known properties of the antenna response we can reduce the number of free 
parameters in our solution. This takes us to the second step of calibration.

@@

After calibration, we form snapshot image cubes (frequency mapping to line-of-sight 
direction). The visibilities are gridded using the primary beam as the gridding kernel, placing 
the data in the holographic frame, which has been shown to be the optimal frame for 
combining images in the sense that it preserves all information for parameter estimation 
\citep{Morales:2009,Bhatnagar:2008}. We use a simulation-based primary beam model for MWA 
antennas developed by \citealt{Sutinjo:2015} 
which incorporates an average embedded element pattern for the dipoles, as well as mutual
coupling between dipoles within an antenna.
At this stage we also average in frequency by 
a factor of two by gridding pairs of frequency channels to the same $(u,v)$ plane, resulting in 
a frequency resolution of 160 kHz.

@@

FHD provides all inputs needed for the \eppsilon package, which include weights cubes (for 
proper accumulation of data), variance cubes (for error propagation), and even/odd 
interleaved data cubes (for an unbiased estimator and a direct measurement of the noise). 
We also retain both East-West and North-South polarizations. At this time we can remap 
our coordinates to cosmological units according to the relationships given in 
\citealt{Morales:2004}, where angular and frequency units $(\theta_x,\theta_y,f)$ map to 
cosmological distance units $(r_x,r_y,r_z)$. Often $r_z$ is referred to as the line of sight 
dimension, and denoted as $r_{||}$. Similarly, $(r_x,r_y)$ is the plane perpendicular to the 
line of sight and is denoted as $\mathbf{r_{\perp}}$.

@@

The top panel of Figure~\ref{fig:cables} shows a 2D power spectrum using the gain model 
described so far. Three of the horizontal lines in the EoR window can be attributed to the 
coarse band gaps, as they reside at the harmonics expected for 1.28~MHz periodic 
sampling. But the sampling cannot account for the fourth line, highlighted by the arrow. The 
corresponding delay time of the line, $\tau \approx 1.23$~$\mu$s, corresponds almost 
exactly to twice the signal travel time through the 150 meter cables (with velocity 0.81c). We 
therefore introduced a reflection term into our gain model, which allows our restricted gains 
to account for the interference of the incident signal with a round-trip reflected signal in the 
cable. 
A similar approach was taken at lower frequencies in \citealt{Ewall-Wice:2016}.
The full gain model is expressed as
\begin{equation}\label{eq:cal2}
\hat{g}_i(\nu)=B_{\alpha}(\nu)\left(P_i(\nu)+R_i(\nu)\right),
\end{equation}
where the reflection term can be further decomposed as
\begin{equation}
R_i(\nu) = R_{i,0} e^{-2\pi i \tau_{i} \nu}.
\end{equation}
The reflection coefficient, $R_{i,0}$ is allowed to be complex, while the reflection delay, 
$\tau_i$ is real. After fitting for all other parameters, we fit for the reflection mode. This is 
done for antennas with suspected reflections, the most offensive of which is seen in the 
antennas with cables of 150 meters. Because the cables were not cut at exact lengths 
(variations on order tens of centimeters), we solve for both the reflection coefficient and 
delay by performing a direct Fourier transform to a highly over resolved delay grid and 
selecting the mode where the reflection amplitude is largest. 
This fitting produces a single reflection coefficient for each antenna with nominal cable
length of 150 meters.

@@

Our pipeline uses two modes of FHD -- the full deconvolution mode to identify point sources 
and build a catalog, and a production mode where we calibrate using the sky model and 
subtract it from the data without further fitting. Because full deconvolution on every 
observation from the MWA is computationally not feasible, we currently restrict ourselves to 
the golden data set to build our model. The process of building the catalog is presented in 
P.~A.~Carroll et al. 2016 (in review). They applied machine learning classification methods 
to select reliable detections from the full deconvolution FHD mode, culminating in the KGS 
catalog. The work here uses an early iteration of the KGS catalog that was available at the 
time of analysis.

@@

Ultimately we need model visibilities to subtract from our data. However, rather than storing 
a model for all observations (different pointings and phase centers), we find it most simple 
to treat the model as a single HEALPix image which can be used to create model visibilities 
for each independent observation. FHD treats each pixel in the image as a point source at 
the pixel location with total flux density equal to the surface brightness of the diffuse image times the area 
of the pixel and creates model visibilities for each snapshot in the same way that it imports 
a catalog of point sources. This is similar to the strategy employed by 
\citealt{Thyagarajan:2015} to model diffuse structure.

@@

The final mask we apply is the wedge. We found that a small buffer beyond the horizon was
necessary to completely mask out the wedge, consistent with \citealt{Dillon:2015}. We
implement the buffer by increasing the slope of the horizon line by 14\%.
This line is shown as the
dashed diagonal line in the left panel of Figure~\ref{fig:slices}.

@@

As demonstrated in \citealt{Jacobs:2016}, comparison between independent 
analysis pipelines is essential to evaluate the validity of algorithms under active development. 
Such comparison provides insight into the strengths and weaknesses of the pipelines, as well 
as confirmation of results. The analysis discussed so far has been based on the FHD to 
\eppsilon pipeline, and here we compare with the Real Time System \citep[RTS,][]
{Mitchell:2008,Ord:2010} to Cosmological \ion{H}{1} Power Spectrum \citep[CHIPS,][]
{Trott:2016} pipeline. Below we present a brief overview of the reference pipeline and 
highlight lessons learned from the comparison.

@@

The source catalog used with RTS was a combined cross-match created with the Positional 
Update and Matching Algorithm (PUMA, J.~L.~B.~Line et al. 2016, in review). Full details of PUMA and 
this cross-matched catalog can be found in Line et al. 2016, but we include a brief 
description here. PUMA is designed to cross-match low radio-frequency ($\lesssim 1\,$GHz) 
catalogs. It utilizes a Bayesian probabilistic positional cross-match approach based on~
\citealt{Budavari:2008}, combined with criteria based on catalog resolution and fitting the 
spectral energy distribution to a power-law. These criteria allow for surveys with differing 
resolutions, which can introduce confusing sources into the matched process.
Primarily the source catalog was based on the MWA Commissioning Survey
\citep[MWACS, ][]{Hurley-Walker:2014}, which was cross-matched to the $74\,$MHz Very 
Large Array Low Frequency Sky Survey redux~\citep[VLSSr, ][]{Lane:2012}, the Molonglo 
Reference Catalogue~\citep[MRC, ][]{Large:1981}, the $843\,$MHz Sydney University 
Molonglo Sky Survey~\citep[SUMSS,][]{Mauch:2003}; and the $1.4\,$GHz NRAO VLA Sky 
Survey~\citep[NVSS,][]{Condon:1998}. MWACS covers approximately 
$20.5^h < \textrm{RA} < 8.5^h$,  $-ˆ’58^\circ < \delta < -ˆ’14^\circ$, so to complete the sky 
coverage needed, MRC was used as a base catalogue outside of the MWACS coverage and 
was matched to VLSSr, SUMSS and NVSS.

@@

\section{Discussion}\label{sec:discussion}
Inspecting our 32 hour integrated power spectra we see two places where we are limited. 
First there is a region that is leakage-dominated at low $k$. The cause of this leakage is yet 
unknown, but likely due to imperfect calibration, beam models, and/or foreground models.
In order to improve on our limit in this regime, future analysis will require calibration which 
better accounts for the instrumental response, perhaps by increasing signal to noise 
through multiple-snapshot solutions or refined antenna response models. 
Through simulation, \citealt{Thyagarajan:2016} recently demonstrated the necessity of
 precise foregrounds and instrumental models in order to retain a clean EoR window.
Studies of the 
beam are underway \citep[e.g.][]{Neben:2015, Sutinjo:2015}, and will likely help to further isolate the 
foreground contaminates in our power spectra. The foreground model is a constant area of 
investigation, but improved spectral dependence in both the point source catalog and the 
diffuse model, as well as the addition of an all-sky Galactic model, will continue to improve 
the foreground subtraction and further unlock the EoR window.

@@

\textbf{Primary beam model.} In order to properly model the instrument response to the sky
we also need an accurate primary beam model. This is especially important for EoR
power spectrum measurements because the (often bright) near-horizon emission resides
at the edge of the foreground wedge 
\citep{Pober:2016, Thyagarajan:2015, Thyagarajan:2015b}, but the low elevation beam
can be difficult to model at the necessary precision \citep{Thyagarajan:2016}. 
\citealt{Asad:2016} demonstrated the need for a polarized primary beam model in order
to mitigate the risk of diffuse polarized emission leaking into Stokes I, conflating with a 
potential cosmological measurement. The MWA
primary beam model is continuously improving to incorporate mutual coupling between the 
phased dipoles, and embedded element patterns for each of the 32 dipoles on each antenna 
\citep{Sutinjo:2015}. Other methods are also being pursued to directly measure the full beam 
response in situ using the ORBCOMM satellite constellation \citep{Neben:2015}, and 
drone-flown calibrator sources (Jacobs, et al. in prep). The latter method is similar to
the strategy \citealt{Virone:2014} used to verify SKA beam patterns.

>>>>

Several probes are being used to investigate the EoR. 
Studies of the polarization of the CMB have placed integrated constraints on 
the timing of reionization \citep[e.g.][]{Planck:2015b, Planck:2015, WMAP9}. Meanwhile, 
observations of highly redshifted quasars have placed upper bounds on redshifts by which 
reionization is complete. For example, \citet{Fan:2006a} showed that reionization must be 
complete by $z \approx 6$, a result which was confirmed independent of model by 
\citet{McGreer:2015}.
Using a $z_{\mathrm{end}}>6$ prior, the most recent analysis of \emph{Planck} data 
found a reionization redshift $z_{\mathrm{re}} = 8.8 \pm 0.9$ for a redshift-symmetric model
($z_{\mathrm{re}} = 8.5 \pm 0.9$ for redshift-asymmetric), and a duration of $\Delta z < 2.8$
\citep{Planck:2016}.
Deep optical and infrared galaxy surveys are also beginning to 
reach the redshifts necessary to further constrain reionization \citep[e.g.][]{Bouwens:2014}, 
and the James Webb Space Telescope will improve on their sensitivity 
\citep{Gardner:2006}.

@@

Using the first generation of instruments, several upper limits have been placed on the 21 
cm EoR signal \citep{Ali:2015, Dillon:2014, Parsons:2014, Jacobs:2015, Paciga:2013}, and 
\citet{Pober:2015} was able to place constraints on physical reionization models. 
However, much work is to be done to understand the data produced by these arrays. 
Foreground subtraction and isolation has emerged as a priority in the field. Central to most 
analysis strategies is the concept of an ``EoR window" -- the region of Fourier space where 
spectrally smooth foregrounds have been isolated from the isotropic (spherically symmetric 
in Fourier space) cosmological signal \citep{Morales:2006, Bowman:2009}. More recent 
studies have shown the existence of a foreground ``wedge", the result of instrumental mode 
mixing, throwing power from spectrally smooth foregrounds to higher Fourier modes 
\citep{Thyagarajan:2015b, Thyagarajan:2015, Trott:2012, Liu:2014a, Liu:2014b, 
Hazelton:2013, Pober:2013, Thyagarajan:2013, Vedantham:2012, Morales:2012, Datta:2010}, 
while others are investigating the spectral structure of point sources themselves \cite[e.g.][]{Offringa:2016}.
The EoR window is still expected to be preserved above the wedge, and 
several analysis pipelines are under active development \citep[e.g. B.~J.~Hazelton et al. 2016, in 
prep; D.~A.~Mitchell et al. 2016, in prep; ][]{Jacobs:2016, Trott:2016, 
Dillon:2015b, Dillon:2013, Trott:2014}.

@@

Our best result is a limit on the cosmological power spectrum of $\Delta^2 \leq 2.7 \times 10^4$ 
mK$^2$ at $k=0.27$ h~Mpc$^{-1}$ and $z=7.1$. This is lower than the GMRT 40 hour
limit ($\Delta^2 \le 6.15 \times 10^4$ mK$^2$ at $z=8.6$; \citealt{Paciga:2013}), but 
significantly higher than the latest PAPER result which 
integrated over 700 hours ($\Delta^2 \le 502$ mK$^2$ at $z=8.4$; \citealt{Ali:2015}).
Several EoR experiments are reaching sensitivity levels where very low systematics
become dominant. The PAPER team saw evidence that antenna cross-talk and foreground
leakage was limiting their result at several scales. The LOFAR team has seen excess noise
and diffuse foreground suppression due to calibration \citep{Patil:2016}. It is essential
for EoR experiments to understand and overcome these systematics in order to reach
a cosmological detection.

@@

\section{The Murchison Widefield Array and Observations}\label{sec:MWA}
The Murchison Widefield Array is one of several first generation radio interferometers with 
a primary science goal of detecting the 21cm EoR power spectrum. The remote Australian 
outback offers relative isolation from human-made radio frequency interference (RFI) such as 
FM radio, or TV stations. A recent study of the RFI environment at the Murchison Radio 
Observatory can be found in \citet{Offringa:2015}.

While the 21 cm EoR signal is priority, the MWA is a general observatory serving several 
science programs beyond cosmology including Galactic and extragalactic surveys, time 
domain astrophysics, solar monitoring, and ionospheric science. The array was thus 
designed with these programs in consideration and the layout was optimized using a 
pseudo-random antenna placement algorithm \citep{Beardsley:2012} to obtain a well 
behaved point spread function while retaining a dense core for EoR sensitivity 
\citep{Beardsley:2013, Bowman:2006}.\footnote{A drone flyover view of the array layout is available on YouTube at \url{https://youtu.be/yDWdTUzTUMo}.}
A full description of the science capabilities of the array is found in
\citet{Bowman:2013}.
The high imaging capability of the MWA enables us to calibrate and subtract foregrounds
based on sky models, leveraging the full capability of the observatory for the
EoR experiment. 
This strategy is in contrast to more targeted EoR experiments like PAPER, 
which utilize a highly redundant layout to enhance sensitivity. Redundant arrays can 
exploit symmetries of the instrument for quick calibration and analysis, but have poor 
point spread functions, making foregrounds more difficult to characterize.

The technical design of the MWA is reviewed in \citet{Tingay:2013}, and we highlight a 
few key characteristics that will become important in our analysis of the data. Each 
antenna of the MWA comprises 16 dual-polarization dipoles placed on a regular grid, lying 
on a ground screen. The radio frequency signals from these dipoles feed into an analog 
beamformer which uses physical delays to ``point" the antenna. The MWA contains 128 of 
these antennas, with a tightly packed core of radius 50 m, and extending out to a radius of
1.5 km for higher resolution calibration and imaging.

@@

\section{Analysis Pipeline}\label{sec:analysis}
In order to ensure consistency in analysis, the international MWA EoR collaboration has 
defined two independent reference analysis pipelines. The details of this strategy are 
outlined in \citet{Jacobs:2016}. This study is based on the FHD-to-\eppsilon 
pipe, which is in turn based on the Fast Holographic Deconvolution 
(FHD\footnote{\url{https://github.com/EoRImaging/FHD}}, \citealt{Sullivan:2012}) and Error 
Propagated Power Spectrum with Interleaved Observed Noise (\eppsilon
\footnote{\url{https://github.com/EoRImaging/eppsilon}}, B.~J.~Hazelton et al. 2016, in 
prep) packages. The general flow of the pipeline is shown in Figure~\ref{fig:pipe}, and we 
give an overview below. As a point of comparison, and to demonstrate robustness of the results, we compute the output power spectra with an independent calibration and power spectrum estimation pipeline, and demonstrate consistency.

@@

Calibration is accomplished within the FHD package in two steps. First we match the raw 
data to model visibilities formed using a realistic simulation of the array response to our 
foreground model. This step results in independent complex gains for each antenna, 
80~kHz frequency channel, and polarization. In the second step we impose restrictions on these 
gains motivated by our understanding of the spectral response of the instrument. By 
reducing the number of free parameters in this step we increase our signal to noise on 
calibration solutions and avoid absorbing unmodeled confusion source flux density into our gain 
solutions which has been shown to contaminate the EoR window \citep{Barry:2016}. 
Recently, \citet{Patil:2016} demonstrated excess noise and a suppression of diffuse
foregrounds in LOFAR images after calibration and subtraction of bright discrete
foreground sources. They suggested a multi-frequency calibration solution as a potential
solution to these systematics, similar to the scheme implemented here. Indeed we have
not seen evidence of these systematics through simulations and direct propagation of
our noise \citep[Hazelton et al. 2016, in prep.]{Barry:2016}.

@@

FHD utilizes the StEFCal algorithm described in \citet{Salvini:2014} to find the minimum 
$\chi^2$ estimate with respect to the complex gains. The result of this operation is 
estimated gains for every antenna, frequency channel, and polarization. However, with 
certain known properties of the antenna response we can reduce the number of free 
parameters in our solution. This takes us to the second step of calibration.

@@

After calibration, we form snapshot image cubes (frequency mapping to line-of-sight 
direction). The visibilities are gridded using the primary beam as the gridding kernel, placing 
the data in the holographic frame, which has been shown to be the optimal frame for 
combining images in the sense that it preserves all information for parameter estimation 
\citep{Morales:2009,Bhatnagar:2008}. We use a simulation-based primary beam model for MWA 
antennas developed by \citet{Sutinjo:2015} 
which incorporates an average embedded element pattern for the dipoles, as well as mutual
coupling between dipoles within an antenna.
At this stage we also average in frequency by 
a factor of two by gridding pairs of frequency channels to the same $(u,v)$ plane, resulting in 
a frequency resolution of 160 kHz.

@@

FHD provides all inputs needed for the \eppsilon package, which include weights cubes (for 
proper accumulation of data), variance cubes (for error propagation), and even/odd 
interleaved data cubes (for an unbiased estimator and a direct measurement of the noise). 
We also retain both East-West and North-South polarizations. At this time we can remap 
our coordinates to cosmological units according to the relationships given in 
\citet{Morales:2004}, where angular and frequency units $(\theta_x,\theta_y,f)$ map to 
cosmological distance units $(r_x,r_y,r_z)$. Often $r_z$ is referred to as the line of sight 
dimension, and denoted as $r_{||}$. Similarly, $(r_x,r_y)$ is the plane perpendicular to the 
line of sight and is denoted as $\mathbf{r_{\perp}}$.

@@

The top panel of Figure~\ref{fig:cables} shows a 2D power spectrum using the gain model 
described so far. Three of the horizontal lines in the EoR window can be attributed to the 
coarse band gaps, as they reside at the harmonics expected for 1.28~MHz periodic 
sampling. But the sampling cannot account for the fourth line, highlighted by the arrow. The 
corresponding delay time of the line, $\tau \approx 1.23$~$\mu$s, corresponds almost 
exactly to twice the signal travel time through the 150 meter cables (with velocity 0.81c). We 
therefore introduced a reflection term into our gain model, which allows our restricted gains 
to account for the interference of the incident signal with a round-trip reflected signal in the 
cable. 
A similar approach was taken at lower frequencies in \citet{Ewall-Wice:2016}.
The full gain model is expressed as
\begin{equation}\label{eq:cal2}
\hat{g}_i(\nu)=B_{\alpha}(\nu)\left(P_i(\nu)+R_i(\nu)\right),
\end{equation}
where the reflection term can be further decomposed as
\begin{equation}
R_i(\nu) = R_{i,0} e^{-2\pi i \tau_{i} \nu}.
\end{equation}
The reflection coefficient, $R_{i,0}$ is allowed to be complex, while the reflection delay, 
$\tau_i$ is real. After fitting for all other parameters, we fit for the reflection mode. This is 
done for antennas with suspected reflections, the most offensive of which is seen in the 
antennas with cables of 150 meters. Because the cables were not cut at exact lengths 
(variations on order tens of centimeters), we solve for both the reflection coefficient and 
delay by performing a direct Fourier transform to a highly over resolved delay grid and 
selecting the mode where the reflection amplitude is largest. 
This fitting produces a single reflection coefficient for each antenna with nominal cable
length of 150 meters.

@@

Our pipeline uses two modes of FHD -- the full deconvolution mode to identify point sources 
and build a catalog, and a production mode where we calibrate using the sky model and 
subtract it from the data without further fitting. Because full deconvolution on every 
observation from the MWA is computationally not feasible, we currently restrict ourselves to 
the golden data set to build our model. The process of building the catalog is presented in 
\citet{Carroll:2016}. They applied machine learning classification methods 
to select reliable detections from the full deconvolution FHD mode, culminating in the KGS 
catalog. The work here uses an early iteration of the KGS catalog that was available at the 
time of analysis.

@@

Ultimately we need model visibilities to subtract from our data. However, rather than storing 
a model for all observations (different pointings and phase centers), we find it most simple 
to treat the model as a single HEALPix image which can be used to create model visibilities 
for each independent observation. FHD treats each pixel in the image as a point source at 
the pixel location with total flux density equal to the surface brightness of the diffuse image times the area 
of the pixel and creates model visibilities for each snapshot in the same way that it imports 
a catalog of point sources. This is similar to the strategy employed by 
\citet{Thyagarajan:2015} to model diffuse structure.

@@

The final mask we apply is the wedge. We found that a small buffer beyond the horizon was
necessary to completely mask out the wedge, consistent with \citet{Dillon:2015}. We
implement the buffer by increasing the slope of the horizon line by 14\%.
This line is shown as the
dashed diagonal line in the left panel of Figure~\ref{fig:slices}.

@@

As demonstrated in \citet{Jacobs:2016}, comparison between independent 
analysis pipelines is essential to evaluate the validity of algorithms under active development. 
Such comparison provides insight into the strengths and weaknesses of the pipelines, as well 
as confirmation of results. The analysis discussed so far has been based on the FHD to 
\eppsilon pipeline, and here we compare with the Real Time System \citep[RTS,][]
{Mitchell:2008,Ord:2010} to Cosmological \ion{H}{1} Power Spectrum \citep[CHIPS,][]
{Trott:2016} pipeline. Below we present a brief overview of the reference pipeline and 
highlight lessons learned from the comparison.

@@

The source catalog used with RTS was a combined cross-match created with the Positional 
Update and Matching Algorithm (PUMA, J.~L.~B.~Line et al. 2016, in review). Full details of PUMA and 
this cross-matched catalog can be found in Line et al. 2016, but we include a brief 
description here. PUMA is designed to cross-match low radio-frequency ($\lesssim 1\,$GHz) 
catalogs. It utilizes a Bayesian probabilistic positional cross-match approach based on~
\citet{Budavari:2008}, combined with criteria based on catalog resolution and fitting the 
spectral energy distribution to a power-law. These criteria allow for surveys with differing 
resolutions, which can introduce confusing sources into the matched process.
Primarily the source catalog was based on the MWA Commissioning Survey
\citep[MWACS, ][]{Hurley-Walker:2014}, which was cross-matched to the $74\,$MHz Very 
Large Array Low Frequency Sky Survey redux~\citep[VLSSr, ][]{Lane:2012}, the Molonglo 
Reference Catalogue~\citep[MRC, ][]{Large:1981}, the $843\,$MHz Sydney University 
Molonglo Sky Survey~\citep[SUMSS,][]{Mauch:2003}; and the $1.4\,$GHz NRAO VLA Sky 
Survey~\citep[NVSS,][]{Condon:1998}. MWACS covers approximately 
$20.5^h < \textrm{RA} < 8.5^h$,  $-ˆ’58^\circ < \delta < -ˆ’14^\circ$, so to complete the sky 
coverage needed, MRC was used as a base catalogue outside of the MWACS coverage and 
was matched to VLSSr, SUMSS and NVSS.

@@

\section{Discussion}\label{sec:discussion}
Inspecting our 32 hour integrated power spectra we see two places where we are limited. 
First there is a region that is leakage-dominated at low $k$. The cause of this leakage is yet 
unknown, but likely due to imperfect calibration, beam models, and/or foreground models.
In order to improve on our limit in this regime, future analysis will require calibration which 
better accounts for the instrumental response, perhaps by increasing signal to noise 
through multiple-snapshot solutions or refined antenna response models. 
Through simulation, \citet{Thyagarajan:2016} recently demonstrated the necessity of
 precise foregrounds and instrumental models in order to retain a clean EoR window.
Studies of the 
beam are underway \citep[e.g.][]{Neben:2015, Sutinjo:2015}, and will likely help to further isolate the 
foreground contaminates in our power spectra. The foreground model is a constant area of 
investigation, but improved spectral dependence in both the point source catalog and the 
diffuse model, as well as the addition of an all-sky Galactic model, will continue to improve 
the foreground subtraction and further unlock the EoR window.

@@

\textbf{Primary beam model.} In order to properly model the instrument response to the sky
we also need an accurate primary beam model. This is especially important for EoR
power spectrum measurements because the (often bright) near-horizon emission resides
at the edge of the foreground wedge 
\citep{Pober:2016, Thyagarajan:2015, Thyagarajan:2015b}, but the low elevation beam
can be difficult to model at the necessary precision \citep{Thyagarajan:2016}. 
\citet{Asad:2016} demonstrated the need for a polarized primary beam model in order
to mitigate the risk of diffuse polarized emission leaking into Stokes I, conflating with a 
potential cosmological measurement. The MWA
primary beam model is continuously improving to incorporate mutual coupling between the 
phased dipoles, and embedded element patterns for each of the 32 dipoles on each antenna 
\citep{Sutinjo:2015}. Other methods are also being pursued to directly measure the full beam 
response in situ using the ORBCOMM satellite constellation \citep{Neben:2015}, and 
drone-flown calibrator sources (Jacobs, et al. in prep). The latter method is similar to
the strategy \citet{Virone:2014} used to verify SKA beam patterns.
